Covariates analysis of MSBB RNASeq data

### 2016-01-25 code/Rmd/MSSM_RNASeq.Rmd
minor changes

tree = hclust(as.dist(t(VOOM.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
tmp = gsub('\\..*','',cols)
WGCNA::plotDendroAndColors(tree, 
                           colors = tmp, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

@@

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL, fig.width=15, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[1]]$plotData
```
### Normalisation (custom/manual design)
Re-perform normalisation with custom design matrix (manually with "TISSUE.Dx","SEX","RACE","PMI" as fixed effects and Donor as random effects)
```{r reperform.normalisation, cache=TRUE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,c("TISSUE.Dx","SEX","RACE","PMI"),drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

@@

```{r residualSigCovars.manual, fig.width=15, fig.height=8}
residualSigCovars[["PC_res"]][[1]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "TISSUE.Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

@@

@@

# Expression
expr = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsembleIDs')
write.table(expr, file='expression.tsv', sep = '\t', row.names=F, quote=F)
obj = File('expression.tsv', 
           name = 'Voom Adjusted Weighted Residual Expression (Tissue.Dx added)', 
           parentId = CODE$properties$id)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(covar, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = CODE$properties$id)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)
```

>>>>

tree = hclust(as.dist(t(VOOM.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
tmp = gsub('\\..*','',cols)
WGCNA::plotDendroAndColors(tree, 
                           colors = tmp, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
dev.off()
```

@@

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL, fig.width=15, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom/manual design)
Re-perform normalisation with custom design matrix (manually with "TISSUE.Dx","SEX","RACE","PMI" as fixed effects and Donor as random effects)
```{r reperform.normalisation, cache=TRUE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,c("TISSUE.Dx","SEX","RACE","PMI"),drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

@@

```{r residualSigCovars.manual, fig.width=15, fig.height=8}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "TISSUE.Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

@@

```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$TISSUE.Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```

@@

# Expression
expr = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsembleIDs')
write.table(expr, file='expression.tsv', sep = '\t', row.names=F, quote=F)
obj = File('expression.tsv', 
           name = 'Voom Adjusted Weighted Residual Expression (Tissue.Dx added)', 
           parentId = CODE$properties$id)
expr_obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(covar, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = CODE$properties$id)
covar_obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r covar_obj$properties$id`    |
|  Residual Expression  |  `r expr_obj$properties$id`     |
